---
title: 'Son Bias in the US: Evidence from Business Names^[The data and scripts used are posted here: https://github.com/soodoku/sonny_side.]'
author: "Walter Guillioli and Gaurav Sood"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

Are American businesses with the word "son(s)" more common than the word "daughter(s)"? To answer the question, we assemble data on business names. In the US, businesses are registered with the state. And to allow businesses to check whether a particular name is taken, etc., each state provides a way to search for business names. But most states do not allow for sophisticated searches or return all the relevant search results. Because of these reasons, we can only estimate a very conservative lower bound. In all, based on data from 40 states, across states, the median ratio of businesses with the word 'son(s)' in their name to the word 'daughter(s)' is 12:1. 

When we split by region, we find that the son to daughter ratio is generally higher in the Northeast and the Midwest compared to the South and the West. Analysis by political preferences leaves that democratic majority states have higher ratios than republican majority states. Lastly, counter-intuitively, correlation with state GDP is weak.

## Data and Methods

In the United States, businesses have to register with a state. Each state provides a website to search for business names. But the functionality of these websites varies a lot. Some states do not return all the search results. Some only allow for particular regex searches---returning all the names that include the search term. We describe the concerns in further detail below, including how we address the challenges:

1. **Search results for son(s) are inflated.**  There are three reasons for that:
	
	* Son is part of many English words, from names such as Jason and Robinson to ordinary English words like mason (which can also be a name). 
	* Son is a Korean name. 
	* Some businesses use the word "son" playfully. For instance, "son" is a homonym of sun, and some people use that to create names like "son of a beach." 

	We address the first issue by cleaning the data using regular expressions, only looking for exact matches of son and sons. We don't formally address the other two concerns as carefully browsing the search results suggests that the concerns are not grave. 

2. **Limits in the number of results shown.** Some states return all the relevant search results, but some states limit the number of search results shown. For example, Alabama only displays the first 1,000 results. The truncated returned set has a significant impact since we only know if a particular search for son(s) has more than 1,000 results, but we don't know how many---it could be 10,000 or 500,000. This limitation means that we can only provide a conservative estimate of the lower bound. 

In all, we were able to acquire data for 36 states. Though not all 50 states were covered, we have a good representation of the United States. The 36 states represent 69.9% of the US population, 71.2% of the US GDP and 71% of the registered businesses. 

We enriched our dataset by acquiring state data from other sources to profile the results. These new attributes include: US Region of the state, US Division of the state, population of the state, GPD of the state, major political party of the state and number of establishments on each state. The sources are identified in the References section of this paper.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load R packages
library(tidyverse)
library(rmarkdown)
library(knitr)
library(kableExtra)
```

```{r data, include=FALSE}
#setwd - remove when we post in github
#setwd("C:/Users/wguil/OneDrive/Documents/GitHub/sonny_side/data/")

#Load CSV file from Excel with data collected for the states
d <- read.csv("../states-work-tracker.csv", stringsAsFactors = FALSE)
str(d)
summary(d)
table(d$Status, useNA = "always") #36 states are done as expected
d <- d[, c(2:7,11)] #drop raw and unneeded cols
head(d)

#Load CSV with region and division for each state
regions <- read.csv("../data/_states/states_region_division.csv", stringsAsFactors = FALSE)
regions <- regions[, -c(2)] #drop state code

#Load CSV with population estimate by state
pop <- read.csv("../data/_states/nst-est2019-alldata.csv", stringsAsFactors = FALSE)
pop <- pop[pop$STATE != 0 & pop$STATE != 72,] #Excluse summary rows and Puerto Rico from file
pop <- pop[, c("NAME","POPESTIMATE2019")] #Keep only columns of interest (state name and 2019 population)
sum(pop$POPESTIMATE2019) #328.2M is estimated population
pop$popestimate2019pct <- pop$POPESTIMATE2019 / sum(pop$POPESTIMATE2019) 

#Load CSV with political party winnining 2016 election
party <- read.csv("../data/_states/presidential_election.csv", stringsAsFactors = FALSE)

#Load CSV with GDP by state
gdp <- read.csv("../data/_states/gdp_by_state.csv", stringsAsFactors = FALSE)
gdp <- gdp[,-c(1)] #remove rank column
gdp$State <- trimws(gdp$State) #remove white space from name
gdp$gdp_pct <- gdp$gdp_pct / 100 #make it as actual pct
#gdp$gdp_mm <- as.integer(gdp$gdp_mm)

#Load US Census data for # of businesses per state
census <- read.csv("../data/_states/us_census_firms_2016.csv", stringsAsFactors = FALSE)
census <- census[census$employment_size == "01:  Total",] #Keep only row with total per area/state
census$establishments_pct <- census$establishments_num / 7757807 #% of establishments in US
census <- census[, c("area", "establishments_num", "establishments_pct")] #Keep only columns of interest

#Merge/Join data to have one master dataset
d <- merge(d, regions, by.x = "Name", by.y = "State", all.x = TRUE)
d <- merge(d, pop, by.x = "Name", by.y = "NAME", all.x = TRUE)
d <- merge(d, party, by.x = "Name", by.y = "State", all.x = TRUE)
d <- merge(d, gdp, by.x = "Name", by.y = "State", all.x = TRUE)
d <- merge(d, census, by.x = "Name", by.y = "area", all.x = TRUE)

#Keep only states with estimate data
d <- d[d$Status %in% c("Done", "Gaurav did"),]
summary(d)

sum(d$popestimate2019pct) #69.9%
sum(d$gdp_pct) #71.2%
sum(d$establishments_pct) #71.1%
```

Table 1 shows a sample of 10 states from the final dataset used for analysis. 

```{r summary, echo=FALSE}
cols2print <- c("Name", "estimate", "Region", "popestimate2019pct", "gdp_pct", "establishments_pct", "presidentialelection2016")
colnames2print <- c("State", "Ratio", "US Region", "Population", "GDP", "Establishments", "Political Party")

kable(d[, cols2print], row.names = FALSE, digits =3, 
     col.names = colnames2print,
     caption = "Summary of the Data",
     format = "latex",
     booktabs = T)
```
## Results
In all, as shown in Figure 1 we find that a conservative estimate of son to daughter ratio is between 2:1 to 72:1 across the 36 states where we have data with a median of 12:1. 

```{r, echo=FALSE}
# Custom ggplot theme
cust_theme <- 
theme_minimal() +
theme(panel.grid.major = element_line(color = "#e1e1e1",  linetype = "dotted"),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom",
      legend.key       = element_blank(),
      legend.key.width = unit(1, "cm"),
      axis.title   = element_text(size = 10, color = "#555555"),
      axis.text    = element_text(size = 7, color = "#555555"),
      axis.title.x = element_text(vjust = 1, margin = margin(10, 0, 0, 0)),
      axis.title.y = element_text(vjust = 1),
      axis.ticks   = element_line(color = "#e1e1e1", linetype = "dotted", size = .1),
      axis.text.x  = element_text(vjust = .3),
      plot.margin = unit(c(.5, .75, .5, .5), "cm"))

ggplot(d, aes(x = estimate, y = reorder(Name, estimate))) +
  cust_theme +
  geom_point(color = "#555555", size = 0.75) +
	#scale_x_continuous(limits = c(0, 75),
	#	breaks = seq(0, 75, 5)) +
	xlab("\nRatio") +
	ylab("") +
	ggtitle("Figure 1: Ratio of company names with the word 'son' to 'daughter'")

```


### Covariates of Son Bias

We now proceed to explore how these results vary by location of the state in the United States, by the state population and political party and by its GDP and number of busines establishments.

a) **Differences by US Region**: When we look at the estimate of son to daughter ratio by Region in USA we see states in the Midwest and Northeast with higher ratios when comparing to the West and particularly the South. Figure 2 shows the biggest gap is Midwest vs South with a median ratio of 24:1 vs 5:1 respectively.

```{r, include=FALSE}
table(d$Region)
aggregate(estimate ~ Region, d, FUN = median)

```

b) **Relationship with the Population size of the State**: When we look at the estimate ratio of son vs daughter with the population fo the state we don't see any relationship between these data points. In fact the correlation if basically zero as seen in Figure 3.

```{r, include=FALSE}
cor(d$POPESTIMATE2019, d$estimate)
cor(d$popestimate2019pct, d$estimate)
```

c) **Relationship with the GDP of the State**: We also looked at how the differences of the ratio of business names using son vs daughter could vary by state as it relates to the size of the state in terms of percentage of the gross domesic product (GDP) of the country. We didn't find any evidence of relationship between these two with a correlation of basically zero. The chart is on the appendix B.

```{r, include=FALSE}
cor(d$gdp_mm, d$estimate)
cor(d$gdp_pct, d$estimate)
```

d) **Relationship with the number of Establishments of the State**: We also obtained data from the census organization from US that offers the number of registered establishments in the USA, this is not necesarily the same as the number of business companies registered per state but since obtaining that exact number wasn't not possible we use this as a proxy. Again, no evidence or correlation is observed. The chart is on the appendix C.

```{r, include=FALSE}
cor(d$establishments_num, d$estimate)
cor(d$establishments_pct, d$estimate)
```

e) **Relationship with the major Political Party of the State**: Finally we looked at the voting data from the 2016 elections and compare how the ratio of son vs daughter is when separating states with a majority of Democratics vs Republicans. The ratio tends to be higher on Democratics states with a median of 17 vs a median of 11 for Republican states - see appendix D.

```{r, echo=FALSE}
ggplot(d, aes(x=Region, y=estimate)) + 
  cust_theme +
  geom_boxplot(color = "#555555") +
  xlab("\nRegion") +
  ylab("\nEstimate of son to daughter ratio") +
  ggtitle("Figure 2: Estimate of son to daughter ratio by US Region") +
  coord_flip()
```

```{r, include=FALSE}
table(d$presidentialelection2016)
aggregate(estimate ~ presidentialelection2016, d, FUN = median)
```

```{r, echo=FALSE}
ggplot(d, aes(x=estimate, y=popestimate2019pct)) + 
  cust_theme +
  geom_point(color = "#555555") +
  labs(title="\nFigure 3: Estimate of son to daughter ratio vs % of US population per state", x="\nEstimate of son to daughter ratio",y="\nPercentage of US population in State\n")
```

### States with number of companies found with word son and daughter and ratio

```{r, echo = F}

cols2print <- c("Name", "estimate", "son", "daughter", "conservative_flag")

colnames2print <- c("State", "Son/Daughter",
                    "Companies w/ 'son'", "Companies w/ 'daughter'", 
                    "Conservative")

kable(d[, cols2print], row.names = FALSE, digits =0, col.names = colnames2print, format = "latex",
     booktabs = T)
```

### Estimate of son to daughter ratio vs GDP % of US

```{r, echo=FALSE}
ggplot(d, aes(x=estimate, y=gdp_pct)) + 
  cust_theme +
  geom_point(color = "#555555") +
  labs(title="\nEstimate of son to daughter ratio vs GDP % of US", x="\nEstimate of son to daughter ratio",y="\nGDP Percentage\n")  
```

### Estimate of son to daughter ratio vs # of Establishments

```{r, echo=FALSE}
ggplot(d, aes(x=estimate, y=establishments_pct)) + 
  cust_theme +
  geom_point(color = "#555555") +
  labs(title="\nEstimate of son to daughter ratio vs # of Establishments", x="\nEstimate of son to daughter ratio",y="\nEstablishments\n")
  
```

### Estimate of son to daughter ratio by mayority Political party (2016 elections)

```{r, echo=FALSE}
ggplot(d, aes(x=presidentialelection2016, y=estimate)) + 
  cust_theme +
  geom_boxplot(color = "#555555") +
  xlab("\nPolitcal Party") +
  ylab("\nEstimate of son to daughter ratio") +
  ggtitle("Estimate of son to daughter ratio by mayority Political party (2016 elections)") +
  coord_flip()

```

## Conclusion
There is clearly an inclination to name businesses including the word son(s) vs daughter(s). We found evidence for 36 states that that a conservative estimate of son to daughter ratio is between 2 to 1 to 72 to 1 across the 36 states where we have data with a median of 12 to 1. Despite not having data for 50 states we feel this is a good representation of the whole country since these 36 states represent 69.9% of the US population, 71.2% of the US GDP and 71.% of the registered establishments. 

Although we didn't find any relatioship with the size of the states in terms of GDP, population or number of establishments we do see some differences across regions and political parties dominating the state. We cannot conclude any causality because of this but further exploration is recommended.

\newpage
## References

1. Halpert, Chris. US census bureau regions and divisions.  
<https://github.com/cphalpert/census-regions/>

2. Kaushik, Saurav. Beginner's Guide on Web Scraping in R (using rvest) with hands-on example.   
<https://bit.ly/2Gj0sF6>

3. Kingl, Arvid. Web Scraping in R: rvest Tutorial.  
<https://www.datacamp.com/community/tutorials/r-web-scraping-rvest>

3. United States Census Bureau. State Population Totals and Components of Change: 2010-2019.  
<https://www.census.gov/data/tables/time-series/demo/popest/2010s-state-total.html>

4. United States Census Bureau. SUSB Historical Data.  
<https://www.census.gov/data/tables/time-series/econ/susb/susb-historical.html>

5. Wikipedia contributors. (2020, January 13). Political party strength in U.S. states. In Wikipedia, The Free Encyclopedia. Retrieved 03:11, January 20, 2020.   
<https://bit.ly/37o6AYB>
